### 2018-01-29 includes/experiment.tex

    We generated several problem suites with \fuzzer{} that made one solver perform poorly, but not others. These suites are \theSuites{}. Figure~\ref{fig:cvc-hard} shows the suites that were uniquely difficult for \cvc{}. Figure~\ref{fig:z3str3-hard} shows the suites that were uniquely difficult for \us{}. All experiments were run in series on the same Linux computer, with a timeout of 15 seconds.

@@

        We describe the defects and missing heuristics exposed by \fuzzer{} to illustrate why \fuzzer{} is a valuable asset in solver developments.

@@

The first issue we found caused \us{} to perform poorly when equivalence classes 
of arithmetic terms were very large (e.g. the lengths of many strings were equal s
to 0). This issue was detected by the \texttt{Concats} problem suite. The slowdown 
was caused by a loop over terms of such equivalence class to search for an 
integer constant. The fix was to check the root of the equivalence class for the 
integer constant, as Z3 makes "interpreted terms" (e.g. constants) the root of 
equivalence classes whenever possible. This issue was last observed in commit 
\texttt{6308636}, and was fixed in commit \texttt{3865c45}.

The second issue, which was exposed by the \texttt{Lengths} suite, caused \us{} to 
perform poorly when a variable has a large length constraint. The slowdown was 
caused by the solver performing a linear search for a satisfying length when 
producing a model for the variable. This issue was an inconsistency between the 
implementation and the behavior described in the \us{} paper \cite{z3str3}. 
This issue was last observed in commit \texttt{66bc68f}, and was fixed in commit 
\texttt{7b536e9}.

@@

Discovering heuristics is crucial for solver efficacy improvements. Take 
conflicting constraint detection as an example. Solvers will eventually catch 
the offending terms, if they follow the string constraint solving algorithms and 
gradually propagate the (dis)equivalences discovered along the way. However, in 
many scenarios, a solver may be able to detect the conflicts before it tries to 
figure out the relationships among sub-terms and actually propagates the facts. 
The benefits are clear. It saves the efforts to reason about sub-terms and 
propagate the observations. As reported in existing works, heuristics usually 
lead to significant improvements. However, this is an extremely challenging 
process as the heuristic are usually learned from examples but it's usually hard 
to having multiple instances confirming such inefficiencies in the first place.

The instances in the \texttt{concats-big} generated by \fuzzer{} helped us 
discover a missing heuristic in \us{}. In particular, \us{} didn't make full use 
of the solving context (e.g. some terms are empty strings) to simplify the 
concatenations of a long list of string terms before trying to reason about the 
equivalences among sub-terms. It hence introduced a large number of unnecessary 
intermediate variables and fact propagations. We partially fixed this issue in 
our develop branch and observed some improvements.

>>>>

    We generated several problem suites with \fuzzer{} that each made a solver perform poorly, but not the others. 
    These suites are \theSuites{}. For example, Figure~\ref{fig:cvc-hard} shows the suites that were uniquely difficult for \cvc{}. 
    Figure~\ref{fig:z3str3-hard} shows the suites that were uniquely difficult for \us{}. 
    We ran all experiments in series on a COMPUTER DESCRIPTION AND EXACT SPECS!!, with a timeout of 15 seconds.

@@

        We illustrate \fuzzer{}'s utility by describing the defects and vulnerable heuristics it exposed.

@@

The first issue we found caused \us{} to perform poorly when equivalence classes 
of arithmetic terms were very large (e.g. the lengths of many strings were equal
to 0\todo{HUH!?}). This issue was detected by the \texttt{Concats} problem suite. The slowdown 
was caused by a loop over terms of such equivalence class to search for an 
integer constant. The fix was to check the root of the equivalence class for the 
integer constant, as Z3 makes "interpreted terms" (e.g. constants) the root of 
equivalence classes whenever possible. This issue was fixed in commit \texttt{3865c45}.

The second issue, which was exposed by the \texttt{Lengths} suite, caused \us{} to 
perform poorly when a variable has a large length constraint. The slowdown was 
caused by the solver performing a linear search for a satisfying length when 
producing a model for the variable. This issue was an inconsistency between the 
implementation and the behavior described in the \us{} paper \cite{z3str3}. 
This issue was fixed in commit 
\texttt{7b536e9}.

@@

Discovering heuristics is crucial for solver efficacy improvements. Take 
conflicting constraint detection as an example\todo{describe what it is. what are offending terms?}. Solvers will eventually catch 
the offending terms, if they follow the string constraint solving algorithms and 
gradually propagate the (dis)equivalences discovered along the way. However, in 
many scenarios, a solver may be able to detect the conflicts before it tries to 
figure out the relationships among sub-terms and actually propagate the facts. 
The benefits are clear. It saves the effort to reason about sub-terms and 
propagate the observations. As reported in existing work, heuristics usually 
lead to significant improvements. However, this is an extremely challenging 
process because it is usually hard to have multiple instances confirming such inefficiencies in the first place.

The instances in the \texttt{concats-big} generated by \fuzzer{} helped us 
discover a vulnerable heuristic in \us{}. In particular, \us{} didn't make full use 
of the solving context (e.g. some terms are empty strings) to simplify the 
concatenations of a long list of string terms before trying to reason about the 
equivalences among sub-terms. It hence introduced a large number of unnecessary 
intermediate variables and fact propagations. We partially fixed this issue in 
our develop branch and observed some improvements.

### 2018-01-29 includes/introduction.tex

    Many SMT solvers supporting the theory of strings (e.g., CVC4 \cite{cvc4},  Z3str2 \cite{z3str2}, Norn \cite{norn} and Z3str3 \cite{z3str3}) have been developed over the past decade. To validate and benchmark their solvers, many authors have relied on hand-crafted input suites \cite{cvc4-tests} \cite{z3str3-tests} \cite{z3str2-tests} or real-world examples from industrial applications \cite{kaluza} \cite{kausler}. These test suites have helped developers identify implementation defects and develop more sophistic solving heuristics. Although extremely helpful, these suites can only cover a limited range of the problem space because they are either manually created or translated from carefully developed programs.

    Fuzzing, on the other hand, has already been widely used in similar scenarios to systematically generate more diverse test cases. Guided by different rules, it can also be applied to generate different instances that are particularly designed to challenge selected capacities. Inspired by traditional fuzzers, we introduce a modular fuzzer for string solvers, \fuzzer{}, and describe its utility as an exploratory validation tool. We demonstrate its effectiveness by presenting defects and limitations we were able to identify in leading solvers.
%     Moreover, a standard input language for problem instances, \smtfull{}\cite{smt}, has been developed and widely adopted.

    \subsection{Contributions}

        \begin{enumerate}
            \item \textbf{The \fuzzer{} tool}: We describe a modular fuzzer which can transform and generate \smtfull{} string and regex problems. We briefly document its components and modular architecture. We provide example usages to demonstrate its usefulness as an exploratory tool for validating solvers. We additionally detail its regex-generating and transforming capabilities.
            \item \textbf{A repository of \smtfull{} problems}: We present a repository of \smtfull{} string and regex problem suites we generated using \fuzzer{}. We demonstrate the usefulness of these suites by revealing specific defects they helped uncover and fix in \us{}.
            \item \textbf{Experimental Results}: We compare \theSolvers{} on the \fuzzer{} suites and selectively present the experimental result on \theSuites{}. We highlighted these suites because they uniquely made some solvers perform poorly, but not others.
            \item \textbf{Analysis}: We analyze our experimental results by inspecting the execution traces of the solvers on the given problems. For \us{}, we pinpoint algorithmic limitations that cause the poor performance. For other solvers, we draw superficial conclusions based solely on the execution traces, and invite their authors to conduct further analysis.
        \end{enumerate}

>>>>

    Many SMT solvers supporting the theory of strings (e.g., CVC4 \cite{cvc4},  Z3str2 \cite{z3str2}, Norn \cite{norn} 
    and Z3str3 \cite{z3str3}) have been developed over the past decade. 
    To validate and benchmark their solvers, many developers have relied on hand-crafted suites \cite{cvc4-tests} 
    \cite{z3str3-tests} \cite{z3str2-tests} or industrial examples \cite{kaluza} \cite{kausler}. 
    These test suites have helped developers identify implementation defects and develop more sophistic solving heuristics. 
    However, the suites cover a limited problem space, can be difficult to create, and hard to acquire.

    In contrast, fuzzing can systematically generate diverse test cases with little effort. 
    Guided by different rules, fuzzing can also be applied to generate instances designed to challenge selected capacities. 
    Inspired by traditional fuzzers, we introduce a fuzzer for string solvers, \fuzzer{}, 
    and describe its utility as an exploratory validation tool. 
    We demonstrate its effectiveness by presenting defects and limitations it helped identify in leading solvers.
    More specifically, we contribute
    \begin{enumerate}
        \item \textbf{The \fuzzer{} tool}. We describe a modular fuzzer which can transform and generate \smtfull{} string and regex problems. We briefly document its components and architecture. 
        We provide examples to demonstrate its usefulness as an exploratory tool for validating solvers. 
        We additionally detail its regex-generating and transforming capabilities.
        \item \textbf{A repository of \smtfull{} problems}. We present a repository of \smtfull{} string and regex problem suites 
        generated by \fuzzer{}. We demonstrate the usefulness of these suites by revealing specific defects they helped uncover and fix in \us{}.
        \item \textbf{Experimental Results and Analysis}. We compare \theSolvers{} on the \fuzzer{} suites and present the experimental results. 
        We highlight some suites that cause some solvers to perform poorly poorly in comparison to the others,
        and inspect solver execution traces for some. In the case of \us{}, we
        pinpoint algorithmic limitations that cause the identified poor performance.
    \end{enumerate}

### 2018-01-29 includes/stringfuzz.tex

This tool transforms \smt{} problems. It supports several transformers and options that specify its output and input, which are explained in Table~\ref{tbl:transformers}. Note that transformers \textit{translate} and \textit{reverse} also preserve problem satisfiability \cite{ifaz}.

@@

This tool takes an \smt{} problem as input and outputs its properties: the number of variables/literals, the max/median syntactic depth of expressions, the max/median literal length, etc.

@@

        \fuzzer{} can be used to generate problems to be saved as static suites. It can also be used in a randomised fashion for exploratory debugging. As an example, the following usage can feed random problems to \cvc{} until it returns an error:
\begin{center}
\vspace{-0.05in}
\begin{varwidth}{\linewidth}
{\scriptsize\begin{verbatim}while stringfuzzg -r random | tee problem.smt25 | cvc4 --lang smt2.5 --tlimit=5000; do
    sleep 0
done
\end{verbatim}}
\end{varwidth}
\end{center}
        Some \unix{} scripting can yield other interesting results. For example, the following script finds a regex problem on which \us{} times out after 5 seconds:
\begin{center}
\vspace{-0.1in}
\begin{varwidth}{\linewidth}
        {\scriptsize\begin{verbatim}while true; do
    stringfuzzg -r regex -c -a increasing -r 2 -t 5 -d 3 > problem.smt25
    result=`z3str3 -T:5 problem.smt25`
    echo $result
    if [ "$result" == "timeout" ]; then
        cat hard-problem.smt2
        break
    fi
done\end{verbatim}}
\end{varwidth}
\end{center}

>>>>

\transformer{} transforms \smt{} problems. Table~\ref{tbl:transformers} lists the several transformers and configurations available.
Note that transformers \textit{translate} and \textit{reverse} preserve problem satisfiability \cite{ifaz} in specific cases.

@@

This tool takes an \smt{} problem as input and outputs a number of properties.
Properties include the number of variables/literals, the max/median syntactic depth of expressions, and the max/median literal length.

@@

        \fuzzer{} can be used to generate problems to be saved as static suites. It can also be used in a randomized fashion for exploratory debugging. As an example, the following usage can feed random problems to \cvc{} until it returns an error:
\begin{center}
\vspace{-0.05in}
\begin{varwidth}{\linewidth}
{\scriptsize\begin{verbatim}while stringfuzzg -r random | tee problem.smt25 | cvc4 --lang smt2.5 --tlimit=5000; do
    sleep 0
done
\end{verbatim}}
\end{varwidth}
\end{center}
        Some \unix{} scripting can yield other interesting results. For example, the following script finds a regex problem on which \us{} times out after 5 seconds:
\begin{center}
\vspace{-0.1in}
\begin{varwidth}{\linewidth}
        {\scriptsize\begin{verbatim}while true; do
    stringfuzzg -r regex -c -a increasing -r 2 -t 5 -d 3 > problem.smt25
    result=`z3str3 -T:5 problem.smt25`
    echo $result
    if [ "$result" == "timeout" ]; then
        cat hard-problem.smt2
        break
    fi
done\end{verbatim}}
\end{varwidth}
\end{center}
